---
title: "Exploratory Data Analysis"
author: "ak5357"
date: "2024-11-14"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(haven)
library(rvest)
library(dplyr)
library(stringr)
library(purrr)

# Do not eval anything, when the file is knit
knitr::opts_chunk$set(eval = FALSE)
```

## Import diabetes data

```{r import_data}
data_xpt = read_xpt("data/raw_data/LLCP2023.XPT")

variables = c("_STATE", "ADDEPEV3", "ASTHMA3", "ASTHNOW", "BPHIGH6", "CADULT1", "CELSXBRT", "CHCCOPD3", "CHCKDNY2", "CHKHEMO3", "COVIDPO1", "CVDCRHD4", "CVDINFR4", "CVDSTRK3", "DIABAGE4", "DIABEDU1", "DIABETE4", "DIABTYPE", "DIFFWALK", "EDUCA", "EXERANY2", "EXEROFT1", "FLUSHOT7", "HAVARTH4", "HEIGHT3", "INSULIN1", "LADULT1", "LNDSXBRT", "PDIABTS1", "PREDIAB2", "PREGNANT", "SMOKE100", "TOLDHI3", "WEIGHT2")

diabetes_df = data_xpt |> 
  select(!!variables) |> 
  janitor::clean_names()
```

```{r export_data}
write_csv(diabetes_df, "data/diabetes_data.csv")
```

## Import metadata

_Note: The output .csv file has a bunch of random instances of the Â character. Ignore these or remove manually._
```{r import_metadata}
# Read in HTML
codebook_html = read_html("data/raw_data/USCODE23_LLCP_091024.HTML")

# Remove all junk tables
metadata_tables = codebook_html |> 
  html_table(header = TRUE) |> 
  keep(~ (nrow(.x) > 0 & ncol(.x) > 1))

# Convert the list of tables into metadata table
metadata_list = metadata_tables |>
  lapply(names) |> 
  map(~ .x[[1]]) |> 
  unlist() |>
  str_replace_all(pattern = "([A-Z0-9])([Q])", replacement = "\\1&&\\2") |> 
  str_replace_all(pattern = "([a-z0-9])([A-Z])", replacement = "\\1&&\\2") |> 
  str_replace_all(pattern = "([\\_|\\.|?|\\:])([Q])", replacement = "\\1&&\\2") |> 
  str_replace_all(pattern = "([\\.|?|\\:])([S])", replacement = "\\1&&\\2") |> 
  str_replace_all(pattern = "MISection", replacement = "MI&&Section") |> 
  str_replace_all(pattern = "DNASection", replacement = "DNA&&Section") |> 
  str_replace_all(pattern = "\\)Section", replacement = "\\)&&Section") |> 
  str_replace_all("Question\\sPrologue:Question", "Question") |> 
  str_split("&&") |> 
  map(~ str_split(.x, ":\\s", n = 2))  # Split by ": " into key-value pairs
  
metadata_list = metadata_list[2:length(metadata_list)]

# Create Metadata Dataframe through for loop
metadata_df = tibble(row_id = 0)

for (i in 1:349){
  # Create temp tibble, new_row
  new_row = tibble(row_id = i)
  
  for (j in 1:(length(metadata_list[[i]]))){
    # Pull out name and value from metadata_list
    name_value_pair = metadata_list[[i]][[j]]
    
    # Colname
    col = name_value_pair[[1]]
    
    if(length(metadata_list[[i]][[j]]) > 1){
      # Value
      val = name_value_pair[[2]]
      
      # Add to new_row
      new_row =
        new_row |> 
        mutate(!!col := val)
    }
  }
  
  # Add new row (temp_tibble) to metadata_df
  metadata_df = metadata_df |> 
    bind_rows(janitor::clean_names(new_row))
}

# Clean metadata df
metadata_df = metadata_df |> 
  slice(-1) |> 
  select(row_id, sas_variable_name, label, question, section_name, type_of_variable) |> 
  filter(sas_variable_name %in% variables) |> 
  arrange(sas_variable_name)
```

```{r export_metadata}
# Save metadata df to csv
write_csv(metadata_df, "data/metadata.csv")
```

To look up metadata of some variable, can use this block. Just change the var_name variable.
```{r var_metadata_search}
var_name = "diab"

metadata_df |> 
  filter(str_detect(sas_variable_name, str_to_upper(var_name))) |> 
  view()

```

# Beginning to write Final Report Deliverable 
Overview: In this project, we analyze Type II Diabetes (T2D) in the U.S through the development of a customizable dashboard that users can use to gain valuable information about their individual risk factors for T2D development based on the metrics they enter.  input specific criteria and our software would yield relevant data. Our group’s inherent interest in public health and chronic disease epidemiology prompts us to explore the context of T2D in this project.

Motivation: T2D is among the leading causes of mortality in the U.S. Through our project, we conduct an in-depth analysis of risk factors and lifestyle choices that contribute to its development. We consider potential the impact of co-variates and co-morbitities through linear and multivariate regression models. .
The deliverable includes graphical representations of regression models, maps of T2D across the US, and a customizable dashboard outlining potential risk factors for T2D. We hope to create a dashboard that provides a predictive estimate to help users understand how they can mitigate their risk and prevent developing T2D in the future. 

Data Source: Our data comes from the Behavioral Risk Factor Surveillance System (BRFSS) [https://www.cdc.gov/brfss/annual_data/annual_2023.html] developed and aggregated by the US Centers for Disease and Prevention (CDC). The data captures diabetes data as related to demographics, individual behaviors, chronic health conditions, health-care access, and use of preventive services. This longitudinal survey has been collecting data since 1984, and was recently updated in September 2024. The data was only available in an 'xpt' format. Hence, to make the data accessible for import and analysis in R, the haven package was used to read the xpt file and then write into an accessible csv format. 
